---
title: "Penguins-lab6"
author: "Andrea Cheung"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. load packages and read in penguins.csv data

Packages: tidyverse, car, oneway test
```{r,include=FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our questions: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, Adelie Penguins near Palmer Archipeligo?

Maybe I can use a one way ANOVA....
- Independent observations
- Equal variance (Levene's test)
- Normality (histograms/qqplots)

```{r}

counts <- with(penguins, table(species,sex)) 

penguins2 <- penguins %>% 
  mutate(
    common_name = case_when(
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
      species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
    )
  ) %>% #simplify names
  select(-species) %>%  #remove a column
  filter(sex == "FEMALE" )
```

Exploratory graphs and Levenne's test for equal variance
```{r}
hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name, scale = "free")
hists
```
```{r}
# make quantile quantile plot
quant <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) + # on the a qq plot, we need to change fill to color. 
  facet_wrap(~ common_name, scale = "free")
quant

#based on qq plots and histograms, observations for each group appear normally distributed. also n > 30 for all groups so means will be normally distributed by CLT. 

```

Evaluate variances and perform a Levene's Test for equal variances
```{r}
# If the largest sample variance is <4x greater than the smallest sample variances, then these are "close enough" to consider variances equal

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances

#The largest variance (33.1) is definitely not more than 4x greater than the smalles vriance (15.2)
```

```{r}
# Levene's Test
# leveneTest function in the 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)

levene_penguin

#H0: No difference in variances (variances are equal)
#HA: variances are NOT equal

#there is no significnat difference in variances across species. 


```

One-way ANOVA for penguin flipper lengths by species (common_name)

```{r}
penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# Pr(>F) is very small. the likelyhood that we got means for flipper length that are as different as our means by random chance is very unlikely. 

```

```{r}
#Post-hoc testing by Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

#aov.test function in the 'onewaytest' packages DOES allow for inline referencing 
```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p <0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05).
